---
title: "HRA-Pop Cell Type Correlation Analysis Notebook"
output: html_document
---

# Environment

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(plyr)
library(knitr)
library(stringr)
library(ggplot2)
library(magrittr)
options(scipen=999)
```
 
# Load Data

```{r data}
# Data set metadata
metadata <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas-ad-hoc/as-datasets-modality.csv",
                     header=T, fileEncoding = "UTF8")

# Data for heatmaps
# Includes organ, anatomical structure, cell types measures by tool and donor gender
data <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-v6.csv", 
                 header=T, fileEncoding = "UTF8")

# Spatial data aggregation: Organs Small & Large Intestines, AS, cell types measures
data_spatial <- read.csv("../heatmaps/spatial/HRApop_Spatial_OASCT_CellCount_Percentage.csv",
                         header=T)

# Cell Type Crosswalks - Data Cleaning
# https://github.com/hubmapconsortium/hra-workflows-runner/tree/main/crosswalking-tables
# New version June 2024
## Azimuth
azimuth <- read.csv("https://cdn.humanatlas.io/digital-objects/ctann/azimuth/v1.0/assets/azimuth-crosswalk.csv",
                header=T, fileEncoding = "UTF8", skip=10)

## Cell Typist
celltypist <- read.csv("https://cdn.humanatlas.io/digital-objects/ctann/celltypist/v1.0/assets/celltypist-crosswalk.csv",
                header=T, fileEncoding = "UTF8", skip=10)

## PopV
popv <- read.csv("https://cdn.humanatlas.io/digital-objects/ctann/popv/v1.0/assets/popv-crosswalk.csv",
                header=T, fileEncoding = "UTF8", skip=10)
```

# Select organs data for heatmaps

```{r select-prep}
#Metadata
metadata$organ_label <- 
  metadata$organ_label %>% 
    str_replace("left ","") %>%
    str_replace("right ","") %>%
    str_replace("Set of lactiferous glands in breast","mammary gland") %>%
    str_replace("respiratory system", "lung") %>%
    str_replace("male reproductive system","prostate") %>%
    str_replace(" of body","") %>%
    str_to_title()

metadata$as_label <- 
  metadata$as_label %>% 
    str_replace("heart ","") %>%
    str_replace(" of body","") %>%
    str_replace("part of ","") %>%
    str_replace("Posteromedial head of posterior papillary muscle of ","") %>%
    str_replace("Right Lateral", "Lateral") %>%
    str_replace("Right Medial", "Medial") %>%
    str_replace("Left apical", "Apical") %>%
    str_replace(" of right mammary gland","") %>%
    str_replace(" of prostate","") %>%
    str_replace(" of spleen","") %>%
    str_replace(" of liver","") %>%
    str_replace(" of kidney","") %>%
    str_replace(" of urinary bladder","") %>%
    str_to_title()

metadata <-
  metadata %>%
  select(modality,tool,organ_label,as_label,dataset) %>% 
  distinct() %>%
  arrange(organ_label,as_label,modality,tool) %>%
  ddply(.(tool,organ_label,as_label), summarise,
        ds_count = length(unique(dataset)))%>%
  arrange(organ_label,tool,ds_count) %>%
  plyr::rename(c("organ_label" = "organ",
                 "as_label" = "as"))

# Save results
write.csv(metadata,"../heatmaps/HRApop_organ_dataset_metadata_pivot.csv",
          row.names = F)

# Calculate summary statistics for metadata analysis
metadata_pivot <- 
  metadata %>%
  ddply(.(organ), summarise,
        Tools = length(unique(tool)),
        AS = length(unique(as)),
        Median_Data_Set_Count = median(ds_count),
        Min_Data_Set_Count = min(ds_count),
        Max_Data_Set_Count = max(ds_count)) %>%
  arrange(desc(AS), desc(Tools), desc(Median_Data_Set_Count), organ) %>%
  plyr::rename(c("organ"="Organ"))

# Save results
metadata_pivot
write.csv(metadata_pivot,"../heatmaps/HRApop_organ_dataset_metadata_pivot2.csv",
          row.names = F)
```

Notes on organs: 

|Organ              |Tools|AS|Min/Max Data Sets|
|-------------------|-----|--|-----------------|
|**Heart**          |    3| 5|            14,64|
|**Large Intestine**|    3| 6|           2-3,11|
|**Small Intestine**|    3| 7|           2-3,10|
|**Lung**           |    3| 4|             3,34|
|Skin               |    3| 1|              2,8|
|Liver              |    2| 3|          2 (all)|
|**Kidney**         | 1 (azi)| 6|         3,117|
|Ureter             | 1 (azi)| 1|       8 (all)|
|Mammary Gland      |1 (popv)| 1|       3 (all)|
|Prostate Gland     |1 (popv)| 2|       4 (all)|
|Spleen             |1 (popv)| 2|          6,11|
|Bladder            |1 (popv)| 1|       7 (all)|

# Prepare Data

```{r data prep}
# Prepare crosswalk field names
names(celltypist) <- names(popv) <- names(azimuth) <- c("organ_level","organ_id",
                                                        "annotation_label","annotation_label_id",
                                                        "cell_label","cell_id","cell_match")

# Title Case for Annotation Labels
popv$annotation_label <- str_to_title(popv$annotation_label)
azimuth$annotation_label <- str_to_title(azimuth$annotation_label)
celltypist$annotation_label <- str_to_title(celltypist$annotation_label)

# Clean up organ label strings
data$sex <- factor(data$sex)

# Prepare Tool
data$tool <- factor(data$tool, 
                  levels=c("azimuth","celltypist","popv","sc_proteomics"),
                  labels=c("Azimuth","CellTypist","popV","SC Proteomics"))

# Prepare organ labels
data$organ <- data$organ %>% 
  str_replace("left ", "") %>%
  str_replace("right ", "") %>%
  str_replace("Set of ", "") %>%
  str_replace("left ", "") %>%
  str_replace(" of body", "") %>% 
  str_replace("respiratory system","lung") %>%
  str_replace("male reproductive system","prostate") %>%
  str_to_title()
data$organ <- factor(data$organ)

# Prepare anatomical structure labels
data$as <- data$as %>% 
  str_replace("heart ","") %>%
  str_replace(" of body","") %>%
  str_replace("part of ","") %>%
  str_replace("Right Lateral", "Lateral") %>%
  str_replace("Right Medial", "Medial") %>%
  str_replace("Left apical", "Apical") %>%
  str_replace(" of right mammary gland","") %>%
  str_replace(" of prostate","") %>%
  str_replace(" of spleen","") %>%
  str_replace(" of liver","") %>%
  str_replace(" of kidney","") %>%
  str_replace(" of urinary bladder","") %>%
  str_to_title()

data$cell_label <- data$cell_label %>%
  str_replace("_"," ")

# Common labels for data frames.
labels <-  c("sex", "tool", "modality", "organ", "organId", 
             "as", "as_id",  "cell_id", "cell_label", "cell_count",
             "percentage", "asct_relation_in_asctb_table",
             "indirect_asct_relation_in_asctb_table", 
             "annotation_label", "annotation_label_id", "cell_label2")
```

## Clean-up Cell Type Names

```{r cell_type_names}
# Create Organ based subsets for cleaning cell type labels
# Heart - 1086
heart <- data[data$organ==levels(data$organ)[1],]
# Lungs - 826
lung <- data[data$organ==levels(data$organ)[6],]
# Kidney - 920
kidney <- data[data$organ==levels(data$organ)[2],]
# Large Intestine - 1790
largeIntestine <- data[data$organ==levels(data$organ)[4],]
# Small Intestine
smallIntestine <- data[data$organ==levels(data$organ)[9],]
# Spleen
spleen <- data[data$organ==levels(data$organ)[10],]
# Prostate
prostate <- data[data$organ==levels(data$organ)[7],]

# Cleaning Azimuth Author Annotations - Heart, Lung, Kidney
# Heart
tmp1 <- 
  join(heart[heart$tool=="Azimuth",c(1:11)],
       azimuth[azimuth$organ_level=="Heart_L2" &
               azimuth$cell_match=="skos:exactMatch" | 
               azimuth$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Lung
tmp2 <- 
  join(lung[lung$tool=="Azimuth",c(1:11)],
       azimuth[azimuth$organ_level=="Lung_v2_finest_level" &
               azimuth$cell_match=="skos:exactMatch" | 
               azimuth$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Kidney
tmp3 <- 
  join(kidney[kidney$tool=="Azimuth",c(1:11)],
       azimuth[azimuth$organ_level=="Kidney_L3" &
               azimuth$cell_match=="skos:exactMatch" | 
               azimuth$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Clean names
 names(tmp1) <- names(tmp2) <- names(tmp3) <- 
   c("sex","tool","modality","organ","organId","as","as_id",
     "cell_id","cell_label","cell_count", "percentage",
     "annotation_label","annotation_label_id","cell_label2")
# Combine Azimuth temp files
data_azimuth <- rbind(tmp1,tmp2,tmp3) 
#Add back missing cell labels when annotation label is missing
data_azimuth[is.na(data_azimuth$annotation_label),]$cell_label2 <- 
  data_azimuth[is.na(data_azimuth$annotation_label),]$cell_label

# Cleaning CellTypist Author Annotations - Heart, Lung, Small & Large Intestine
# Heart
tmp1 <- 
  join(heart[heart$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="Healthy_Adult_Heart_pkl" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Lung
tmp2 <- 
  join(lung[lung$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="Human_Lung_Atlas_pkl" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Large Intestines
tmp3 <- 
  join(largeIntestine[largeIntestine$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="intestine_L1" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Small Intestines
tmp4 <- 
  join(smallIntestine[smallIntestine$tool=="CellTypist",c(1:11)],
       celltypist[celltypist$organ_level=="intestine_L1" &
                  celltypist$cell_match=="skos:exactMatch" | 
                  celltypist$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Clean names
 names(tmp1) <- names(tmp2) <- names(tmp3) <- names(tmp4) <-
   c("sex","tool","modality","organ","organId","as","as_id",
     "cell_id","cell_label","cell_count", "percentage",
     "annotation_label","annotation_label_id","cell_label2")
# Combine Azimuth temp files
data_celltypist <- rbind(tmp1,tmp2,tmp3,tmp4)
#Add back missing cell labels when annotation label is missing
data_celltypist[is.na(data_celltypist$annotation_label),]$cell_label2 <- 
  data_celltypist[is.na(data_celltypist$annotation_label),]$cell_label

# Cleaning popV Author Annotations - Heart, Lung, Small & Large Intestine
# Heart
tmp1 <- 
  join(heart[heart$tool=="popV",c(1:11)],
       popv[popv$organ_level=="heart" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left") 
# Lung 
tmp2 <- 
  join(lung[lung$tool=="popV",c(1:11)],
       popv[popv$organ_level=="lung" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Large Intestines
tmp3 <- 
  join(largeIntestine[largeIntestine$tool=="popV",c(1:11)],
       popv[popv$organ_level=="large intestine" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Small Intestines
tmp4 <- 
  join(smallIntestine[smallIntestine$tool=="popV",c(1:11)],
       popv[popv$organ_level=="small intestine" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Spleen
tmp5 <- 
  join(spleen[spleen$tool=="popV",c(1:11)],
       popv[popv$organ_level=="spleen" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")
# Prostate
tmp6 <- 
  join(prostate[prostate$tool=="popV",c(1:11)],
       popv[popv$organ_level=="prostate gland" &
            popv$cell_match=="skos:exactMatch" | 
            popv$cell_match=="skos:narrowMatch", c(3:6)], 
       by="cell_id", type="left")

# Clean names
names(tmp1) <- names(tmp2) <- names(tmp3) <- 
  names(tmp4) <- names(tmp5) <- names(tmp6) <- 
  c("sex","tool","modality","organ","organId","as","as_id",
    "cell_id","cell_label","cell_count", "percentage",
    "annotation_label","annotation_label_id","cell_label2")
# Combine Azimuth temp files
data_popv <- rbind(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6) 
#Add back missing cell labels when annotation label is missing
data_popv[is.na(data_popv$annotation_label),]$cell_label2 <- 
  data_popv[is.na(data_popv$annotation_label),]$cell_label


# SC Proteomics
# Large Intestine 
tmp1 <- 
  largeIntestine[largeIntestine$tool=="SC Proteomics",c(1:11)] %>%
  mutate(annotation_label=NA,
         annotation_label_ids=NA,
         cell_label2=NA)
# Small Intestine 
tmp2 <- 
  smallIntestine[smallIntestine$tool=="SC Proteomics",c(1:11)] %>%
  mutate(annotation_label=NA,
         annotation_label_ids=NA,
         cell_label2=NA)
# Clean names
names(tmp1) <- names(tmp2) <-
   c("sex","tool","modality","organ","organId","as","as_id",
     "cell_id","cell_label","cell_count", "percentage",
     "annotation_label","annotation_label_id","cell_label2")
# Combine Azimuth temp files
data_scproteomics <- rbind(tmp1,tmp2) 
#Add back missing cell labels when annotation label is missing
data_scproteomics[is.na(data_scproteomics$annotation_label),]$cell_label2 <- 
  data_scproteomics[is.na(data_scproteomics$annotation_label),]$cell_label

# Add cell_name field to CTann data subsets.
data_popv$cell_name <- 
  paste0(data_popv$cell_id,": ",data_popv$cell_label2)
data_celltypist$cell_name <- 
  paste0(data_celltypist$cell_id,": ",data_celltypist$cell_label2)
data_azimuth$cell_name <- 
  paste0(data_azimuth$cell_id,": ",data_azimuth$cell_label2)
data_scproteomics$cell_name <- 
  paste0("ASCTB-TEMP: ",data_scproteomics$cell_label2)

# Spatial cell_label updated to cell_name. Underlying data did not include AS_ID
names(data_spatial)[4] <- "cell_name"

# Clean up environment
rm(tmp1,tmp2,tmp3,tmp4,tmp5,tmp6)
rm(heart,lung,kidney,largeIntestine,smallIntestine)
rm(azimuth,popv,celltypist)
```

# Heatmaps Analysis

## Azimuth

```{r heatmap_azimuth_v0.2, fig.height=12, fig.width=24}
# Create pivot
data_azimuth <- 
  data_azimuth %>%
  select(organ, as, as_id, cell_name, percentage) %>%
  pivot_wider(names_from = cell_name,
              values_from = percentage,
              values_fill = 0,
              values_fn = mean) %>%
  arrange(organ, as_id)

# Convert DF to a matrix
data_azimuth_heatmap <- data.matrix(data_azimuth[,4:ncol(data_azimuth)])

# Assign Row.names as organs
rownames(data_azimuth_heatmap) <- 
  paste0(data_azimuth$organ,"-",data_azimuth$as)
data_azimuth_heatmap <- 
  data_azimuth_heatmap[order(row.names(data_azimuth_heatmap)),]

# Scale data
data_azimuth_heatmap <- scale(data_azimuth_heatmap)

# Generate heatmap
heatmap_plot <- pheatmap::pheatmap(data_azimuth_heatmap, legend=TRUE,
                                   legend_breaks = c(-1, 0, 1, 2, 3, 
                                                     max(data_azimuth_heatmap)),
                                   legend_labels = c("-1", "0", "1", "2", "3", 
                                     "Percentage (z-scale)\n"))

# Save visualization
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-azimuth.pdf"),
       heatmap_plot, width=24, height=12, units="in")
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-azimuth.png"),
       heatmap_plot, width=24, height=12, units="in")

# Plot
heatmap_plot
```

## CellTypist

```{r heatmap_celltypist_v0.2, fig.height=12, fig.width=36}
# Create pivot
data_celltypist <- 
  data_celltypist %>% 
  select(organ, as, as_id, cell_name, percentage) %>%
  pivot_wider(names_from = cell_name,
              values_from = percentage,
              values_fill = 0,
              values_fn = mean) %>%
  arrange(organ, as_id)

# Convert DF to a matrix
data_celltypist_heatmap <- data.matrix(data_celltypist[,4:ncol(data_celltypist)])

# Assign Row.names as organs
rownames(data_celltypist_heatmap) <- 
  paste0(data_celltypist$organ,"-",data_celltypist$as)
data_celltypist_heatmap <- 
  data_celltypist_heatmap[order(row.names(data_celltypist_heatmap)),]

# Scale data
data_celltypist_heatmap <- scale(data_celltypist_heatmap)

# Generate heatmap
heatmap_plot <- pheatmap::pheatmap(data_celltypist_heatmap, legend=TRUE,
                                   legend_breaks = c(-1,0, 1, 2, 3, 4,
                                                     max(data_celltypist_heatmap)),
                                   legend_labels = c("-1","0", "1", "2", "3", "4",
                                     "Percentage (z-scale)\n"))

# Save visualization
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-celltypist.pdf"),
       heatmap_plot, width=36, height=12, units="in")
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-celltypist.png"),
       heatmap_plot, width=36, height=12, units="in")

# Plot
heatmap_plot
```

## PopV

```{r heatmap_popV_v0.2, fig.height=12, fig.width=24}
# Create pivot
data_popv <- 
  data_popv %>%
  select(organ, as, as_id, cell_name, percentage) %>%
  pivot_wider(names_from = cell_name,
              values_from = percentage,
              values_fill = 0,
              values_fn = mean) %>%
  arrange(organ, as_id)

# Convert DF to a matrix
data_popv_heatmap <- data.matrix(data_popv[,4:ncol(data_popv)])

# Assign Row.names as organs
rownames(data_popv_heatmap) <- 
  paste0(data_popv$organ,"-",data_popv$as)
data_popv_heatmap <- 
  data_popv_heatmap[order(row.names(data_popv_heatmap)),]

# Scale data
data_popv_heatmap <- scale(data_popv_heatmap)

# Generate heatmap
heatmap_plot <- pheatmap::pheatmap(data_popv_heatmap, legend=TRUE,
                                   legend_breaks = c(-1, 0, 1, 2, 3, 4,
                                                     max(data_popv_heatmap)),
                                   legend_labels = c("-1", "0", "1", "2", "3", "4",
                                     "Percentage (z-scale)\n"))

# Save visualization
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-popv.pdf"),
       heatmap_plot, width=24, height=12, units="in")
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-popv.png"),
       heatmap_plot, width=24, height=12, units="in")

# Plot
heatmap_plot
```


## SC Proteomics

```{r heatmap_scproteomics_v0.1, fig.height=6.5, fig.width=12}
# Create pivot
data_scproteomics <- 
  data_scproteomics %>%
  select(organ, as, as_id, cell_name, percentage) %>%
  pivot_wider(names_from = cell_name,
              values_from = percentage,
              values_fill = 0,
              values_fn = mean) %>%
  arrange(organ, as_id)

# Convert DF to a matrix
data_scproteomics_heatmap <- data.matrix(data_scproteomics[,4:ncol(data_scproteomics)])

# Assign Row.names as organs
rownames(data_scproteomics_heatmap) <- 
  paste0(data_scproteomics$organ,"-",data_scproteomics$as)
data_scproteomics_heatmap <- 
  data_scproteomics_heatmap[order(row.names(data_scproteomics_heatmap)),]

# Scale data
data_scproteomics_heatmap <- scale(data_scproteomics_heatmap)

# Generate heatmap
heatmap_plot <- pheatmap::pheatmap(data_scproteomics_heatmap, legend=TRUE,
                                  legend_breaks = c(-2, -1, 0, 1, 2,
                                                    max(data_scproteomics_heatmap)),
                                  legend_labels = c(-2, -1, 0, 1, 2,
                                                    "Percentage (z-scale)\n"))

# Save visualization
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-scproteomics.pdf"),
       heatmap_plot, width=12, height=6.5, units="in")
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-scproteomics.png"),
       heatmap_plot, width=12, height=6.5, units="in")

# Plot
heatmap_plot
```

## Spatial

```{r heatmap_spatial_v0.1, fig.height=5, fig.width=12}
names(data_spatial) <- c("organ","as_id","as","cell_name","datasets",
                         "cell_count","mean_cell_ct","total_cell_ct",
                         "percentage","tool")

# Create pivot
data_spatial <- 
  data_spatial %>%
  select(organ, as, as_id, cell_name, percentage) %>%
  pivot_wider(names_from = cell_name,
              values_from = percentage,
              values_fill = 0,
              values_fn = sum) %>%
  arrange(organ, as_id)

# Convert DF to a matrix
data_spatial_heatmap <- data.matrix(data_spatial[,4:ncol(data_spatial)])

# Assign Row.names as organs
rownames(data_spatial_heatmap) <- 
  paste0(data_spatial$organ,"-",data_spatial$as)
data_spatial_heatmap <- 
  data_spatial_heatmap[order(row.names(data_spatial_heatmap)),]

# Scale data
data_spatial_heatmap <- scale(data_spatial_heatmap)

# Generate heatmap
heatmap_plot <- pheatmap::pheatmap(data_spatial_heatmap, legend=TRUE,
                                   legend_breaks = c(-1, 0, 1, 2,
                                                     max(data_spatial_heatmap)),
                                   legend_labels = c("-1", "0", "1", "2",
                                                     "Percentage (z-scale)\n"))

# Save visualization
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-spatial.pdf"),
       heatmap_plot, width=12, height=5, units="in")
ggsave(file=paste0("../output/heatmaps/HRApop_organ_as_celltype-heatmap-spatial.png"),
       heatmap_plot, width=12, height=5, units="in")

# Plot
heatmap_plot
```


