---
title: "A notebook to compile counts for the HRApop paper"
output: html_notebook
---

# Import libraries
```{r }
#libraries
library(tidyverse)
library(rmarkdown) # for rendering this
library(glue)
```

# Set global variables
```{r}
# global variables
hra_pop_version = "0.10.2"
```

# Load data
```{r warning=FALSE, message=FALSE}
sankey_universe = read_csv(paste("../../hra-pop/output-data/v",hra_pop_version,"/reports/universe-ad-hoc/sankey.csv", sep = ""))

# View first few rows
glimpse(sankey_universe)
```
# Pre-processing steps
## Simplify rows with multiple annotations for getting accurate counts
```{r}
# replace tool with sc-transcriptomics, no cell summary, sc-proteomics
tool_replacement = "sc_transcriptomics with Cell Summary"

subset_sankey$cell_type_annotation_tool[subset_sankey$cell_type_annotation_tool == "azimuth"] <- tool_replacement
subset_sankey$cell_type_annotation_tool[subset_sankey$cell_type_annotation_tool == "celltypist"] <- tool_replacement
subset_sankey$cell_type_annotation_tool[subset_sankey$cell_type_annotation_tool == "popv"] <- tool_replacement

duplicates <- subset_sankey[duplicated(subset_sankey),]
duplicates

subset_sankey <- distinct(subset_sankey)
print(subset_sankey)

```

## Manually fix missing cell type annotation values for SenNet atlas datasets
See GitHub issue: https://github.com/x-atlas-consortia/hra-pop/issues/91
``` {r}

sankey_universe$cell_type_annotation_tool[sankey_universe$portal == "SenNet" & sankey_universe$is_atlas_dataset == TRUE] <- tool_replacement
```

# Get counts for HRApop paper
The following sections provide counts of datasets and other metrics for HRApop v0.10.2.

## Report numbers for Highlights
```{r non-atlas datasets highlights}
# all datasets downloaded
all = sankey_universe$unique_dataset_id %>% unique() %>% length()
print(glue("Number of UNIVERSE datasets:: {all}"))

# all sc-proteomics
all_sc_proteomics = sankey_universe %>% select(dataset_id, cell_type_annotation_tool) %>% filter(cell_type_annotation_tool == "sc_proteomics") %>% distinct() %>% count()
print(glue("Number of sc-proteomics datasets: {all_sc_proteomics}"))

# datasets with cell summary 
sc_transcriptomics_with_cell_summary = subset_sankey %>% filter(cell_type_annotation_tool != "No Cell Summary", cell_type_annotation_tool != "sc_proteomics") %>% .$unique_dataset_id %>% length()
print(glue("Number of sc-transcriptomics datasets with cell summary: {sc_transcriptomics_with_cell_summary}"))
```

## Report numbers for Sankey/experimental data
```{r non-atlas datasets experimental data}
# atlas datasets
atlas = sankey_universe %>% filter(is_atlas_dataset == TRUE) 
print(glue("Atlas datasets: {atlas$unique_dataset_id %>% unique() %>% length()}")) 

# datasets with extraction site but without cell summary
no_ctann = sankey_universe %>% filter(is_rui_registered==TRUE, is.na(cell_type_annotation_tool))
print(glue("Datasets with extraction site but without cell summary: {no_ctann$unique_dataset_id %>% unique() %>% length()}"))

# datasets with cell summary but without extraction site
no_rui = sankey_universe %>% filter(is_rui_registered==FALSE, !is.na(cell_type_annotation_tool))
print(glue("Datasets with cell summary but without extraction site: {no_rui$unique_dataset_id %>% unique() %>% length()}")) 

# datasets with neither
neither = sankey_universe %>% filter(is_rui_registered == FALSE, is.na(cell_type_annotation_tool))
print(glue("Datasets with neither: {neither$unique_dataset_id %>% unique() %>% length()}")) 

# number of non-atlas datasets
not_atlas = sankey_universe %>% filter(is_atlas_dataset == FALSE)
print(glue("Non-atlas datasets total: {not_atlas$unique_dataset_id %>% unique() %>% length()}")) 

```