---
title: "Anatomical Structures - CT-Pop Cell Types Stacked Bar Graphs - Notebook"
output:
  # html_document:
  # df_print: paged
  pdf_document: 
    dev: pdf
    fig_caption: no
    number_sections: yes
    toc: yes
    toc_depth: 3
documentclass: article
urlcolor: blue
classoption: a4paper
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
---

# Environment 

```{r environment, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)

options(scipen = 999)

knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align='center',
    fig.pos='H',
    fig.path = "../output/barplots/",
    dev = c("pdf"),
    dpi=500
)
```

# Load Data

```{r load}
# Data - bar graph 
# Includes organ, anatomical structure, cell types measures by tool and donor gender
df <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-v6.csv", 
                 header=T, fileEncoding = "UTF8")

# # Cell Type Crosswalks - Data Cleaning
# # https://github.com/hubmapconsortium/hra-workflows-runner/tree/main/crosswalking-tables
# # New version June 2024
# ## Azimuth
# azi <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/azimuth.csv",
#                 header=T, fileEncoding = "UTF8", skip=10)
# 
# ## Cell Typist
# ctp <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/celltypist.csv",
#                 header=T, fileEncoding = "UTF8", skip=10)
# 
# ## PopV
# pop <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/popv.csv",
#                 header=T, fileEncoding = "UTF8", skip=10)
```

# Preparing Data

```{r prep}
# Prepare crosswalk field names
# names(ctp) <- names(pop) <- names(azi) <- c("organ_level","organ_id",
#                                             "annotation_label","annotation_label_id",
#                                             "cell_label","cell_id","cell_match")
names(df)[5] <- "organ_id"

# Add record number to data 
df$rec <- as.numeric(row.names(df))

# Prepare Sex
df$sex <- factor(df$sex)

# Prepare Tool
df$tool <- factor(df$tool, 
                  levels=c("azimuth","celltypist","popv","sc_proteomics"),
                  labels=c("Azimuth","CellTypist","popV","SC Proteomics"))

# Prepare organ labels
df$organ <- df$organ %>% 
  str_replace("left ", "") %>%
  str_replace("right ", "") %>%
  str_replace("Set of ", "") %>%
  str_replace("left ", "") %>%
  str_replace(" of body", "") %>% 
  str_replace("respiratory system","lung") %>%
  str_replace("male reproductive system","prostate") %>%
  str_to_title()
df$organ <- factor(df$organ)

# Prepare anatomical structure labels
df$as <- df$as %>% 
  str_to_title() %>%
  str_replace(" Part "," part ") %>%
  str_replace(" Of "," of ") %>% 
  str_replace(" In "," in ") %>%
  str_replace("Heart ","")

df$cell_label <- df$cell_label %>%
  str_replace("_"," ")
```


# Subsetting data by Organ

```{r subs}
# Heart - 730 records
hrt <- df[df$organ==levels(df$organ)[1],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)# %>%
  # ddply(.(sex,tool,as,cell_lab), summarise,
  #       ct = length(cell_count),
  #       mean_cell_ct = mean(cell_count),
  #       mean_cell_pt = mean(percentage)) %>%
  # arrange(sex, tool, as, desc(mean_cell_pt))
hrt$as <- factor(hrt$as, 
                 levels=c("Right Ventricle","Left Ventricle",
                          "Posteromedial Head of Posterior Papillary Muscle Of Left Ventricle",
                          "Interventricular Septum","Right Cardiac Atrium","Left Cardiac Atrium"),
                 labels=c("Right Ventricle","Left Ventricle","Left Ventricle",
                          "Interventricular Septum","Right Cardiac Atrium","Left Cardiac Atrium"))
hrt$cell_lab <- factor(hrt$cell_lab)

# Kidney - 898 records - 449 row agg.
kid <- df[df$organ==levels(df$organ)[2],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) %>%
  ddply(.(sex,tool,as,cell_lab), summarise,
        ct = length(cell_count),
        mean_cell_ct = mean(cell_count),
        mean_cell_pt = mean(percentage)) %>%
  arrange(sex, tool, as, desc(mean_cell_pt)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) %>%
  filter(rank<=10)
kid$as <- factor(kid$as)
kid$cell_lab <- factor(kid$cell_lab, levels=c("ASCTB-TEMP:descending-vasa-recta-endothelial: Descending Vasa Recta Endothelial ",
                                              "CL:1001131: Ascending Vasa Recta Endothelial",
                                              "CL:0000890: M2 Macrophage",
                                              "CL:1000452: Parietal Epithelial",
                                              "CL:4030009: Proximal Tubule Epithelial Segment 1",
                                              "CL:4030011: Proximal Tubule Epithelial Segment 3",
                                              "CL:1000692: Fibroblast",
                                              "CL:4030022: Medullary Fibroblast",
                                              "CL:1000714: Cortical Collecting Duct Principal",
                                              "CL:1000715: Cortical Collecting Duct Intercalated Type A",
                                              "CL:1000716: Outer Medullary Collecting Duct Principal",
                                              "CL:1000718: Inner Medullary Collecting Duct",
                                              "CL:1000768: Connecting Tubule",
                                              "CL:4030016: Distal Convoluted Tubule Type 1",
                                              "CL:4030018: Connecting Tubule Principal",
                                              "CL:1001107: Ascending Thin Limb",
                                              "CL:1001108: Medullary Thick Ascending Limb",
                                              "CL:1001109: Cortical Thick Ascending Limb",
                                              "CL:4030012: Descending Thin Limb Type 1",
                                              "CL:4030013: Descending Thin Limb Type 2",
                                              "CL:4030014: Descending Thin Limb Type 3",
                                              "CL:1001318: Vascular Smooth Muscle / Pericyte"))

# Breast - 14 records
brt <- df[df$organ==levels(df$organ)[3],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
brt$as <- factor(brt$as)
brt$cell_lab <- factor(brt$cell_lab)

# Large Intestine - 1714 records
lrg <- df[df$organ==levels(df$organ)[4],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) # %>%
  # ddply(.(sex,tool,as,cell_lab), summarise,
  #       ct = length(cell_count),
  #       mean_cell_ct = mean(cell_count),
  #       mean_cell_pt = mean(percentage)) %>%
  # arrange(sex, tool, as, desc(mean_cell_pt))
lrg$as <- factor(lrg$as)
lrg$cell_lab <- factor(lrg$cell_lab)

# Liver - 81 records
lvr <- df[df$organ==levels(df$organ)[5],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) 
lvr$as <- factor(lvr$as)
lvr$cell_lab <- factor(lvr$cell_lab)

# Lung - 826 records
lng <- df[df$organ==levels(df$organ)[6],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) #%>%
  # ddply(.(sex,tool,as,cell_lab), summarise,
  #       ct = length(cell_count),
  #       mean_cell_ct = mean(cell_count),
  #       mean_cell_pt = mean(percentage)) %>%
  # arrange(sex, tool, as, desc(mean_cell_pt))
lng$as <- factor(lng$as)
lng$cell_lab <- factor(lng$cell_lab)

# Prostate - 24 records
pro <- df[df$organ==levels(df$organ)[7],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
pro$as <- factor(pro$as)
pro$cell_lab <- factor(pro$cell_lab)

# Skin - 50 records
skn <- df[df$organ==levels(df$organ)[8],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
skn$as <- factor(skn$as)
skn$cell_lab <- factor(skn$cell_lab)

# Small Intestine - 2012 records
sml <- df[df$organ==levels(df$organ)[9],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) #%>%
  # ddply(.(sex,tool,as,cell_lab), summarise,
  #       ct = length(cell_count),
  #       mean_cell_ct = mean(cell_count),
  #       mean_cell_pt = mean(percentage)) %>%
  # arrange(sex, tool, as, desc(mean_cell_pt))
sml$as <- factor(sml$as)
sml$cell_lab <- factor(sml$cell_lab)

# Spleen - 75 records
spl <- df[df$organ==levels(df$organ)[10],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) 
spl$as <- factor(spl$as)
spl$cell_lab <- factor(spl$cell_lab)

# Ureter - 71 records
utr <- df[df$organ==levels(df$organ)[11],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
utr$as <- factor(utr$as)
utr$cell_lab <- factor(utr$cell_lab)

# Bladder - 26 records
bld <- df[df$organ==levels(df$organ)[12],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) 
bld$as <- factor(bld$as)
bld$cell_lab <- factor(bld$cell_lab)

```

# Visualization

Base design for 100% stacked bar graphs representing the cell types commonly identified in anatomical structures within donated organs.

For each set of records associated with organs: 
Bars represent an anatomical structure, 
Bar segments are associated with Cell Type percentage measures
   values segments should total to at most 1 or 100%.
   mean cell type percentage calculation - organ, sex, tool, as
Bar color represents cell types
Facets represent combination of analytic tool and gender OR analytic method.

```{r bar_kidney_as_1, fig.height=6, fig.width=20}
color_pal_kid <- c("#CD8490","#A26771","#C0BCC8","#AADCDF","#88B1B3","#556E6F",
                   "#8DC599","#6F9A78","#BAD2E7","#9AADBE","#7E8E9C","#626F7A",
                   "#D6B6D7","#AB92AC","#675868","#E97B74","#E07770","#BE655F",
                   "#9C534E","#7A413D","#582F2C","#A1ACD2")

color_pal_kid2 <- c("#FF0043","#D8D7DB","#70A5A8","#CD8490","#7495AE","#AADCDF",
"#E9E4F3","#D6B6D7","#BAD2E7","#8DC599","#F9CE8D","#EDB8AC","#A295C9","#A1ACD2",
"#4B4B5E","#E97B74","#C4C5C6","#A2A3A4","#818182","#5F6060","#3D3E3E","#cdcecf")


kid %>% arrange(sex,tool,as,mean_cell_pt) %>%
        ggplot(aes(x=as,y=mean_cell_pt)) +
          geom_bar(aes(fill=cell_lab), stat="identity") +
          facet_wrap(facets=vars(tool, sex)) +
          labs(x="Anatomical Structure", y="Percentage (%)") +
          scale_fill_manual(name="Cell Type",values=color_pal_kid) +
          scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
          theme_light() +
          theme(legend.position = "bottom",
                legend.title = element_text(face="bold", size=10),
                strip.background = element_rect(fill = "#cdcecf"),
                strip.text = element_text(color="black", size=10, hjust=0))
```



