---
title: "Anatomical Structures - CT-Pop Cell Types Stacked Bar Graphs - Notebook"
output: html_notebook
---

### Environment 

```{r environment, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(stringr)

options(scipen = 999)

knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align='center',
    fig.pos='H',
    fig.path = "../output/barplots/",
    dev = c("pdf"),
    dpi=500
)
```

### Load Data

```{r load, echo=TRUE}
# Data - bar graph 
# Includes organ, anatomical structure, cell types measures by tool and donor gender
df <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-v6.csv", 
                 header=T, fileEncoding = "UTF8")

# Cell Type Crosswalks - Data Cleaning
# https://github.com/hubmapconsortium/hra-workflows-runner/tree/main/crosswalking-tables
# New version June 2024
## Azimuth
azi <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/azimuth.csv",
                header=T, fileEncoding = "UTF8", skip=10)

## Cell Typist
ctp <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/celltypist.csv",
                header=T, fileEncoding = "UTF8", skip=10)

## PopV
pop <- read.csv("https://raw.githubusercontent.com/hubmapconsortium/hra-workflows-runner/main/crosswalking-tables/popv.csv",
                header=T, fileEncoding = "UTF8", skip=10)
```

### Preparing Data

```{r prep, echo=TRUE}
# Prepare crosswalk field names
names(ctp) <- names(pop) <- names(azi) <- c("organ_level","organ_id",
                                            "annotation_label","annotation_label_id",
                                            "cell_label","cell_id","cell_match")

# Title Case for Annotation Labels
pop$annotation_label <- str_to_title(pop$annotation_label)
azi$annotation_label <- str_to_title(azi$annotation_label)
ctp$annotation_label <- str_to_title(ctp$annotation_label)

# Add record number to data 
df$rec <- as.numeric(row.names(df))

# Prepare Sex
df$sex <- factor(df$sex)

# Prepare Tool
df$tool <- factor(df$tool, 
                  levels=c("azimuth","celltypist","popv","sc_proteomics"),
                  labels=c("Azimuth","CellTypist","popV","SC Proteomics"))

# Prepare organ labels
df$organ <- df$organ %>% 
  str_replace("left ", "") %>%
  str_replace("right ", "") %>%
  str_replace("Set of ", "") %>%
  str_replace("left ", "") %>%
  str_replace(" of body", "") %>% 
  str_replace("respiratory system","lung") %>%
  str_replace("male reproductive system","prostate") %>%
  str_to_title()
df$organ <- factor(df$organ)

# Prepare anatomical structure labels
df$as <- df$as %>% 
  str_to_title() %>%
  str_replace(" Part "," part ") %>%
  str_replace(" Of "," of ") %>% 
  str_replace(" In "," in ") %>%
  str_replace("Heart ","")

df$cell_label <- df$cell_label %>%
  str_replace("_"," ")
```

#### Clean up Cell Labeling

```{r cell_labs}
# Creating Organ Subset for Cell Label clean-up

# Heart - 1086 records
hrt <- df[df$organ==levels(df$organ)[1],]
tmp1 <- join(hrt[hrt$tool=="Azimuth",c(1:13)],
             azi[azi$organ_level=="Heart_L2" &
                 azi$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left")
tmp2 <- join(hrt[hrt$tool=="popV",c(1:13)],
             pop[pop$organ_level=="heart" &
                 pop$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left") 
tmp3 <- join(hrt[hrt$tool=="CellTypist",c(1:13)],
             ctp[ctp$organ_level=="heart_L1" &
                 ctp$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left")
hrt <- rbind(tmp1,tmp2,tmp3) 
hrt[is.na(hrt$annotation_label),]$annotation_label <- 
  hrt[is.na(hrt$annotation_label),]$cell_label
names(hrt)[16] <- "cell_label2"

# Lung - 826 records
lng <- df[df$organ==levels(df$organ)[6],]
tmp1 <- join(lng[lng$tool=="Azimuth",c(1:13)],
             azi[azi$organ_level=="Lung_v2_L5" &
                 azi$cell_match=="skos:exactMatch", c(3:6)],
             by="cell_id", type="left")
tmp2 <- join(lng[lng$tool=="popV",c(1:13)],
             pop[pop$organ_level=="lung" &
                 pop$cell_match=="skos:exactMatch", c(3:6)],
             by="cell_id", type="left")
tmp3 <- join(lng[lng$tool=="CellTypist",c(1:13)],
             ctp[ctp$organ_level=="Human_Lung_Atlas_pkl" &
                 ctp$cell_match=="skos:exactMatch", c(3:6)],
             by="cell_id", type="left")
lng <- rbind(tmp1,tmp2,tmp3)
names(lng)[16] <- "cell_label2"
lng[is.na(lng$annotation_label),]$annotation_label <-
  lng[is.na(lng$annotation_label),]$cell_label


# Kidney - 898 records
kid <- df[df$organ==levels(df$organ)[2],]
kid <- join(kid[kid$tool=="Azimuth",c(1:13)],
             azi[azi$organ_level=="Kidney_L3" &
                 azi$cell_match=="skos:exactMatch", c(3:6)],
             by="cell_id", type="left")

names(kid)[16] <- "cell_label2"
kid[is.na(kid$annotation_label),]$annotation_label <-
  kid[is.na(kid$annotation_label),]$cell_label

# Large Intestine - 1714 records
lrg <- df[df$organ==levels(df$organ)[4],]
tmp1 <- join(lrg[lrg$tool=="SC Proteomics",c(1:13)],
             pop[pop$organ_level=="large intestine" &
                 pop$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left")
tmp2 <- join(lrg[lrg$tool=="popV",c(1:13)],
             pop[pop$organ_level=="large intestine" &
                 pop$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left") 
tmp3 <- join(lrg[lrg$tool=="CellTypist",c(1:13)],
             ctp[ctp$organ_level=="intestine_L1" &
                 ctp$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left")
lrg <- rbind(tmp1,tmp2,tmp3) 
names(lrg)[16] <- "cell_label2"
lrg[is.na(lrg$annotation_label),]$annotation_label <-
  lrg[is.na(lrg$annotation_label),]$cell_label

# Small Intestine - 2012 records
sml <- df[df$organ==levels(df$organ)[9],] 
tmp1 <- join(sml[sml$tool=="SC Proteomics",c(1:13)],
             pop[pop$organ_level=="small intestine" &
                 pop$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left")
tmp2 <- join(sml[sml$tool=="popV",c(1:13)],
             pop[pop$organ_level=="small intestine" &
                 pop$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left") 
tmp3 <- join(sml[sml$tool=="CellTypist",c(1:13)],
             ctp[ctp$organ_level=="intestine_L1" &
                 ctp$cell_match=="skos:exactMatch", c(3:6)], 
             by="cell_id", type="left")
sml <- rbind(tmp1,tmp2,tmp3) 
names(sml)[16] <- "cell_label2"
sml[is.na(sml$annotation_label),]$annotation_label <-
  sml[is.na(sml$annotation_label),]$cell_label

# Clean up
rm(tmp1,tmp2,tmp3)

# Data with clean cell lables for pivots
df2 <- rbind(hrt,lng,kid,lrg,sml) %>%
  select(tool, modality, sex, organ, organId, as, as_id, cell_id, cell_label, 
         annotation_label, cell_count, percentage, 
         asct_relation_in_asctb_table, 
         indirect_asct_relation_in_asctb_table )
# Cell label corrections
df2$annotation_label <- df2$annotation_label %>%
                          str_replace(" Cells", "") %>%
                          str_replace(" Cell", "") %>%
                          str_replace("_cell", "") %>%
                          str_replace(" cell", "") %>%
                          str_replace("Cardiomyocytes", "Cardiomyocyte") %>%
                          str_replace("Cardiomycoyte", "Cardiomyocyte") %>%
                          str_replace("Adipocytes", "Adipocyte") %>%
                          str_replace("Ilc", "ILC")

```

#### Pivot - Identification of Common Cell Types

The goal of this pivot table is to generate descriptive statistics for cell types associated with anatomical structures. Data is aggregated to identify cell types associated with given organ anatomical structures, and the tools that generated the cell type analysis results. Cell type statistics include, the count of anatomical structures related to a cell type, the mean & median of a cell type percentage measure.

```{r piv-vt}
# Pivot 1 - Count Records
piv1 <- df2 %>% 
  filter(asct_relation_in_asctb_table == 'true' |
         indirect_asct_relation_in_asctb_table == 'true') %>%
  select(tool, organ, as, cell_id, cell_label, annotation_label, 
         cell_count, percentage) %>% 
  #mutate(as = factor(as)) %>%
  ddply(.(tool, organ, as, cell_id, annotation_label), summarise,
        measure_ct = length(cell_count),
        mean_cell_ct = mean(cell_count),
        median_cell_ct = median(cell_count),
        mean_cell_per = mean(percentage),
        median_cell_per = median(percentage))
# Pivot 2 - Count AS
piv2 <- piv1 %>% 
  ddply(.(tool, organ, cell_id, annotation_label), summarise,
        as_ct = length(as),
        measure_ct = sum(measure_ct),
        mean_cell_ct = mean(mean_cell_ct),
        median_cell_ct = median(median_cell_ct),
        mean_cell_per = mean(mean_cell_per),
        median_cell_per = median(median_cell_per)) %>% 
  arrange(tool, desc(median_cell_per)) %>%
  group_by(organ, tool) %>%
  mutate(rank = rank(-median_cell_per))

# Pivot 2 - Count Organ
piv3 <- piv2 %>%  
  ddply(.(tool, cell_id, annotation_label), summarise,
        og_ct = length(organ),
        as_ct = sum(as_ct),
        measure_ct = sum(measure_ct),
        mean_cell_ct = mean(mean_cell_ct),
        median_cell_ct = median(median_cell_ct),
        mean_cell_per = mean(mean_cell_per),
        median_cell_per = median(median_cell_per)) %>%
  arrange(tool, desc(og_ct), desc(median_cell_per)) %>%
  group_by(tool) %>%
  mutate(rank = rank(-median_cell_per))

# Create filters of common cell types to exclude from analysis
cell_filter <- piv3 %>%
  ungroup() %>%
  filter(og_ct >= 2) %>%
  select(cell_id) %>%
  distinct() 
```

### Visualization

Base design for 100% stacked bar graphs representing the cell types commonly identified in anatomical structures within donated organs.

Bar graph definition:

Stacked bar represent top cell types associated with an organ anatomical structure
* Data threshold: 8% minimum cell percentage values (or mean cell percentage values PER AS)
* Bars - anatomical structure categories.
  * Bar segment-area-Cell Type percentage measures
    * Bar/Bar segments should total to at most 1 or 100%.
    * Mean cell type percentage calculation-tool, sex organ, as
  * Bar segment-color-cell types categories.
* Facets represent combination of analytic tool and gender.

#### Heart - Anatomical Structure - Cell Types

The code chunk documents the general process used to prepare data for visualizations. In the first step, we subset data related to an single organ. Next, assign the best cell type name associated with an organ, based on the cell type annotation tool used. These tool subsets are re-combined and clean cell label is created. After cleaning the cell identifiers, the data is aggregated to determine the mean count and percentage of a cell types associated with tissue registration data, based on donor sex, analysis tools, and anatomical structure of an organ. After aggregation, cell types that make up at least 8% of cell type populations are selected for visualization. 

```{r bar_heart_as_1, echo=TRUE, fig.height=12, fig.width=14, message=FALSE, warning=FALSE}

hrt <- df2[df2$organ==levels(df$organ)[1],]

hrt <- hrt %>%
  filter(!cell_id %in% cell_filter$cell_id) %>%
  select(organ,as,cell_id,cell_label,annotation_label,
         cell_count,percentage,tool,sex) %>%
  mutate(cell_lab = paste0(cell_id,": ",annotation_label)) %>%
  ddply(.(sex,tool,as,cell_lab), summarise,
        obvs = length(cell_lab),
        cell_count = sum(cell_count),
        mn_pct = mean(percentage)) %>%
  arrange(sex, tool, as, desc(mn_pct)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) %>%
  ungroup()
  
hrt2 <- hrt %>%  
  filter(mn_pct >= .02)

# Factoring AS and Cell Types for heart
hrt2$as <- factor(hrt2$as, 
                 levels=c("Left Ventricle","Right Ventricle",
                          "Posteromedial Head of Posterior Papillary Muscle Of Left Ventricle",
                          "Interventricular Septum","Left Cardiac Atrium","Right Cardiac Atrium"),
                 labels=c("Left Ventricle","Right Ventricle","Left Ventricle",
                          "Interventricular Septum","Left Cardiac Atrium","Right Cardiac Atrium"))

hrt2$cell_lab <- factor(hrt2$cell_lab,
                        levels=c("CL:0000115: EC10 CMC-like",
                                "CL:0000115: EC4 immune",
                                "CL:0000136: Adipocyte",                   
                                "CL:0000182: Hepatocyte",
                                "CL:0000623: Nk",
                                "CL:0000669: PC1 vent",
                                "CL:0000669: PC2 atria",
                                "CL:0000669: PC3 str",
                                "CL:0000669: Pericyte",
                                "CL:0000746: Cardiac Muscle",
                                "CL:0000359: Basic Smc",
                                "CL:0002129: Atrial Cardiomyocyte",
                                "CL:0002131: Ventricular Cardiomyocyte",
                                "CL:0002350: EC7 endocardial",
                                "CL:0002350: Endocardial",
                                "CL:0002396: CD16+Mo",
                                "CL:0002548: Fibroblast Of Cardiac Tissue",
                                "CL:0002573: Neuronal",
                                "CL:0010008: Cardiac Endothelial",
                                "CL:0002144: Capillary Endothelial",
                                "CL:1000413: Arterial Endothelial",
                                "CL:1000414: Venous Endothelial",
                                "CL:0002543: Venous Endothelial",
                                "CL:0010020: NC1 glial",
                                "CL:0011031: Monocyte/Cdc"),
                         labels=c("CL:0000115: EC10 CMC-like",
                                "CL:0000115: EC4 immune",
                                "CL:0000136: Adipocyte",                   
                                "CL:0000182: Hepatocyte",
                                "CL:0000623: Nk",
                                "CL:0000669: Pericyte",
                                "CL:0000669: Pericyte",
                                "CL:0000669: Pericyte",
                                "CL:0000669: Pericyte",
                                "CL:0000746: Cardiac Muscle",
                                "CL:0000359: Basic Smc",
                                "CL:0002129: Atrial Cardiomyocyte",
                                "CL:0002131: Ventricular Cardiomyocyte",
                                "CL:0002350: Endocardial",
                                "CL:0002350: Endocardial",
                                "CL:0002396: CD16+Mo",
                                "CL:0002548: Fibroblast Of Cardiac Tissue",
                                "CL:0002573: Neuronal",
                                "CL:0010008: Cardiac Endothelial",
                                "CL:0002144: Capillary Endothelial",
                                "CL:1000413: Arterial Endothelial",
                                "CL:1000414: Venous Endothelial",
                                "CL:1000414: Venous Endothelial",
                                "CL:0010020: NC1 glial",
                                "CL:0011031: Monocyte/Cdc"))

# Visualization
color_pal_kid2 <- c("#80CBC4","#B2DFDB","#FFD54F","#FFF176","#D4E157","#9CCC65",
                    "#C62828","#E57373","#81D4FA","#03A9F4","#FF8A65","#A1887F",
                    "#A5D6A7","#4CAF50","#E1BEE7","#CE93D8","#BA68C8","#9C27B0",
                    "#B0BEC5","#7986CB","#C4C5C6","#A2A3A4","#818182","#5F6060",
                    "#3D3E3E","#cdcecf")


hrt_bg <- hrt2 %>% arrange(sex,tool,as,mn_pct) %>%
              ggplot(aes(x=as,y=mn_pct)) +
                geom_bar(aes(fill=cell_lab), stat="identity") +
                facet_wrap(facets=vars(tool, sex), ncol=2) +
                labs(x="Anatomical Structure", y="Percentage (%)",
                     title="Heart Anatomical Structures - Cell Types greater than 8% share, by Donor Sex and Analysis Tool") +
                scale_fill_manual(name="Cell Type",values=color_pal_kid2) +
                scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
                theme_light() +
                theme(legend.position = "bottom",
                      legend.title = element_text(face="bold", size=10),
                      legend.key.spacing.y = unit(.3,"pt"),
                      strip.background = element_rect(fill = "#cdcecf"),
                      strip.text = element_text(color="black", size=10, hjust=0))

# Export pdf
ggsave(file=paste0("../output/barplots/heart_as_celltype_bargraph_0.2.pdf"),
       hrt_bg, width=14, height=12, units="in")

hrt_bg
```

#### Lung - Anatomical Structure - Cell Types

```{r bar_lung_1, fig.height=12, fig.width=14, message=FALSE, warning=FALSE, paged.print=FALSE}

lng <- lng %>%
  select(organ,as,cell_id,cell_label,#annotation_label,
         cell_count,percentage,tool,sex) %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  # mutate(cell_lab = paste0(cell_id,": ",annotation_label)) %>%
  # ddply(.(sex,tool,as,cell_lab), summarise,
  #     obvs = length(cell_lab),
  #     cell_count = sum(cell_count),
  #     mn_pct = mean(percentage)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) %>%
  arrange(sex, tool, as, desc(percentage)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) %>%
  filter(percentage >= .08)

# Factor Lung AS
lng$as <- factor(lng$as, 
                 levels=c("Lateral Bronchopulmonary Segment",
                          "Lateral Segmental Bronchus",
                          "Left Apical Bronchopulmonary Segment",
                          "Medial Bronchopulmonary Segment"),
                 labels=c("Lateral\nBronchopulmonary\nSegment",
                          "Lateral\nSegmental\nBronchus",
                          "Left Apical\nBronchopulmonary\nSegment",
                          "Medial\nBronchopulmonary\nSegment"))

# lng$cell_lab <- factor(lng$cell_lab)
# View(as.data.frame(levels(lng$cell_lab)))

# Factor Lung Cell Types
lng$cell_lab <- factor(lng$cell_lab,
                       levels=c("CL:0000057: fibroblast",
                                "CL:4028004: Alveolar fibroblasts",
                                "CL:0000235: macrophage",
                                "CL:0000583: Alveolar macrophages",
                                "CL:0000492: CD4 T cells",
                                "CL:0000624: CD4-positive, alpha-beta T cell",
                                "CL:0002062: AT1",
                                "CL:0002062: type I pneumocyte",
                                "CL:0002063: AT2",
                                "CL:0002063: type II pneumocyte",
                                "CL:0002144: capillary endothelial cell",
                                "CL:0002144: EC general capillary",
                                "CL:0002633: Basal resting",
                                "CL:0005012: Multiciliated (non-nasal)",
                                "CL:1001603: Monocyte-derived Mph",
                                "CL:0000875: Non-classical monocytes"),
                       labels=c("CL:0000057: Fibroblast",
                                "CL:4028004: Alveolar Fibroblasts",
                                "CL:0000235: Macrophage",
                                "CL:0000583: Alveolar Macrophages",
                                "CL:0000492: CD4 T Cells",
                                "CL:0000624: CD4-positive, alpha-beta T Cell",
                                "CL:0002062: AT1",
                                "CL:0002062: Type I pneumocyte",
                                "CL:0002063: AT2",
                                "CL:0002063: Type II pneumocyte",
                                "CL:0002144: Capillary Endothelial Cell",
                                "CL:0002144: EC General Capillary",
                                "CL:0002633: Basal Resting",
                                "CL:0005012: Multiciliated (non-nasal)",
                                "CL:1001603: Monocyte-derived Mph",
                                "CL:0000875: Non-classical Monocytes"))
"CL:0000115: EC10 CMC-like",
"CL:0000115: EC4 immune",
"CL:0000136: Adipocyte",
"CL:0000182: Hepatocyte",
"CL:0000623: Nk",                         
"CL:0000669: PC1 vent",
"CL:0000669: PC2 atria",
"CL:0000669: PC3 str",
"CL:0000669: Pericyte",
"CL:0000746: Cardiac Muscle",
"CL:0000359: Basic Smc",
"CL:0002129: Atrial Cardiomyocyte",
"CL:0002131: Ventricular Cardiomycoyte",
"CL:0002131: Ventricular Cardiomyocyte",
"CL:0002144: Capillary Endothelial",
"CL:0002350: EC7 endocardial",
"CL:0002350: Endocardial",
"CL:0002396: CD16+Mo",                    
"CL:0002548: Fibroblast Of Cardiac Tissue",
"CL:0002573: Neuronal",                 
"CL:0010020: NC1 glial",
"CL:0011031: Monocyte/Cdc",
"CL:0010008: Cardiac Endothelial",
"CL:1000413: Arterial Endothelial",
"CL:0002543: Venous Endothelial",
"CL:1000414: Venous Endothelial"          



color_pal_kid <- c("#CD8490","#AADCDF","#88B1B3",
                   "#BAD2E7","#9AADBE","#7E8E9C","#626F7A",
                   "#D6B6D7","#AB92AC",
                   "#9C534E","#582FC5","#A1ACD2")

color_pal_kid2 <- c("#F9CE8D","#E97B74","#556E6F","#70A5A8","#EDB8AC","#CD8490",
                    "#E9E4F3","#D6B6D7","#A295C9","#AB92AC","#8DC599","#6F9A78",
                    "#9C534E","#C39A5D","#A17F4D","#675868")

lng_bg <- lng %>% arrange(sex,tool,as,percentage) %>%
        ggplot(aes(x=as,y=percentage)) +
          geom_bar(aes(fill=cell_lab), stat="identity") +
          facet_wrap(facets=vars(tool, sex), ncol=2) +
          labs(x="Anatomical Structure", y="Percentage (%)",
               title="Lung Anatomical Structures - Cell Types greater than 8% share, by Donor Sex and Analysis Tool") +
          scale_fill_manual(name="Cell Type",values=color_pal_kid2) +
          scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
          theme_light() +
          theme(legend.position = "bottom",
                legend.title = element_text(face="bold", size=10),
                legend.key.spacing.y = unit(.3,"pt"),
                strip.background = element_rect(fill = "#cdcecf"),
                strip.text = element_text(color="black", size=10, hjust=0)) +
          guides(fill=guide_legend(ncol=4))

# Export pdf
ggsave(file=paste0("../output/barplots/lung_as_celltype_bargraph_0.2.pdf"),
       lng_bg, width=14, height=12, units="in")

lng_bg
```


#### Kidney - Anatomical Structure - Cell Types

```{r bar_kidney_as_1, fig.height=6, fig.width=14}


kid <- kid %>%
  select(organ,as,cell_id,annotation_label,
         cell_count,percentage,tool,sex) %>%
  mutate(cell_lab = paste0(cell_id,": ",annotation_label)) %>%
  ddply(.(sex,tool,as,cell_lab), summarise,
        ct = length(cell_count),
        mean_cell_ct = mean(cell_count),
        mean_cell_pt = mean(percentage)) %>%
  arrange(sex, tool, as, desc(mean_cell_pt)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) %>%
  filter(mean_cell_pt >= .08)

# Factor Kidney AS
kid$as <- factor(kid$as)

# Factor Kidney Cell Types
#View(as.data.frame(levels(factor(kid$cell_lab))))
kid$cell_lab <- factor(kid$cell_lab, 
                       levels=c("CL:1000714: Cortical Collecting Duct Principal",
                                "CL:1000716: Outer Medullary Collecting Duct Principal",
                                "CL:1000718: Inner Medullary Collecting Duct",
                                "CL:1001108: Medullary Thick Ascending Limb",
                                "CL:1001109: Cortical Thick Ascending Limb",
                                "CL:1001107: Ascending Thin Limb",
                                "CL:4030012: Descending Thin Limb Type 1",
                                "CL:4030014: Descending Thin Limb Type 3",
                                "CL:1000768: Connecting Tubule",
                                "CL:4030016: Distal Convoluted Tubule Type 1",
                                "CL:4030009: Proximal Tubule Epithelial Segment 1",
                                "CL:4030022: Medullary Fibroblast"))



color_pal_kid <- c("#CD8490","#A26771","#C0BCC8","#AADCDF","#88B1B3","#556E6F",
                   "#8DC599","#6F9A78","#BAD2E7","#9AADBE","#7E8E9C","#626F7A",
                   "#D6B6D7","#AB92AC","#675868","#E97B74","#E07770","#BE655F",
                   "#7A413D","#582F2C","#A1ACD2","#E9E4F3","#CD8490")

color_pal_kid2 <- c("#7495AF","#70A5A8","#AADCDF","#D6B6D7","#A295C9",
                    "#9C534E","#E97B74","#EDB8AC","#8DC599","#6F9A78","#F9CE8D","#A1ACD2",
                    "#556E6F","#FF0043","#BAD2E7","#D8D7DB","#4B4B5E","#C4C5C6",
                    "#A2A3A4","#818182","#5F6060","#3D3E3E","#cdcecf")

kid_bg <- kid %>% arrange(sex,tool,as,mean_cell_pt) %>%
              ggplot(aes(x=as,y=mean_cell_pt)) +
                geom_bar(aes(fill=cell_lab), stat="identity") +
                facet_wrap(facets=vars(tool, sex)) +
                labs(x="Anatomical Structure", y="Percentage (%)",
                     title="Kidney Anatomical Structures - Cell Types greater than 8% share, by Donor Sex and Analysis Tool") +
                scale_fill_manual(name="Cell Type",values=color_pal_kid2) +
                scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
                theme_light() +
                theme(legend.position = "bottom",
                      legend.title = element_text(face="bold", size=10),
                     # legend.key.spacing.y = unit(.1)
                      strip.background = element_rect(fill = "#cdcecf"),
                      strip.text = element_text(color="black", size=10, hjust=0))

# Export pdf
ggsave(file=paste0("../output/barplots/kidney_as_celltype_bargraph_0.2.pdf"),
       kid_bg, width=14, height=6, units="in")

kid_bg
```

#### Large Intestine - Anatomical Structure - Cell Types

```{r bar_largeIntestine_as_1, fig.height=12, fig.width=14, message=FALSE, warning=FALSE, paged.print=FALSE}

  
# Select data for visualization
lrg <- lrg %>%
  mutate(cell_lab = paste0(cell_id,": ",annotation_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) %>%
  ddply(.(sex,tool,as,cell_lab), summarise,
        obvs = length(cell_lab),
        cell_count = sum(cell_count),
        mn_pct = mean(percentage)) %>%
  arrange(sex, tool, as, desc(mn_pct)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) %>%
  filter(mn_pct >= .08)

# Factor Organ AS  
lrg$as <- factor(lrg$as, levels=c("Caecum","Ascending Colon","Transverse Colon",
                                  "Descending Colon","Sigmoid Colon","Rectum"))
# Factor Cell Types
lrg$cell_lab <- factor(lrg$cell_lab, 
                       levels=c("ASCTB-TEMP:colonocyte: Colonocyte",
                                "ASCTB-TEMP:cycling-plasma-cell: Cycling plasma cell",
                                "ASCTB-TEMP:cycling-ta: Cycling TA",
                                "CL:0000131: Gut Endothelial Cell",
                                "ASCTB-TEMP:endothelial: Endothelial",
                                "CL:0002071: Enterocyte Of Epithelium Of Large Intestine",
                                "ASCTB-TEMP:enterocyte: Enterocyte",
                                "CL:1000320: Large Intestine Goblet Cell",
                                "ASCTB-TEMP:goblet-cell: Goblet cell",
                                "ASCTB-TEMP:goblet: Goblet",
                                "ASCTB-TEMP:m2-macrophage: M2 Macrophage",
                                "ASCTB-TEMP:mesoderm-1-hand1-: Mesoderm 1 (HAND1+)",
                                "ASCTB-TEMP:smooth-muscle: Smooth muscle",
                                "ASCTB-TEMP:smc-plpp2-: SMC (PLPP2+)",
                                "ASCTB-TEMP:stroma: Stroma",
                                "CL:0000057: Fibroblast",
                                "CL:0019032: Intestinal Tuft Cell"),
                        labels=c("ASCTB-TEMP: Colonocyte",
                                "ASCTB-TEMP: Cycling plasma cell",
                                "ASCTB-TEMP: Cycling TA",
                                "CL:0000131: Gut Endothelial Cell",
                                "ASCTB-TEMP: Endothelial",
                                "CL:0002071: Enterocyte Of Epithelium",
                                "ASCTB-TEMP: Enterocyte",
                                "CL:1000320: Goblet cell",
                                "ASCTB-TEMP: Goblet cell",
                                "ASCTB-TEMP: Goblet cell",
                                "ASCTB-TEMP: Macrophage",
                                "ASCTB-TEMP: Mesoderm 1 (HAND1+)",
                                "ASCTB-TEMP: Smooth muscle",
                                "ASCTB-TEMP: SMC (PLPP2+)",
                                "ASCTB-TEMP: Stroma",
                                "CL:0000057: Fibroblast",
                                "CL:0019032: Intestinal Tuft cell")
                       )

color_pal_kid <- c("#CD8490","#AADCDF","#88B1B3",
                   "#BAD2E7","#9AADBE","#7E8E9C","#626F7A",
                   "#D6B6D7","#AB92AC",
                   "#9C534E","#582FC5","#A1ACD2")

color_pal_kid2 <- c("#F9CE8D","#556E6F","#70A5A8","#E9E4F3","#D6B6D7","#EDB8AC","#CD8490",
                    "#AADCDF","#BAD2E7","#7495AF","#A295C9","#8DC599","#6F9A78","#E97B74",
                    "#9C534E","#C39A5D","#A17F4D","#675868")

lrg_bg <- lrg %>% arrange(sex,tool,as,mn_pct) %>%
              ggplot(aes(x=as,y=mn_pct)) +
                geom_bar(aes(fill=cell_lab), stat="identity") +
                facet_wrap(facets=vars(tool, sex), ncol=2) +
                labs(x="Anatomical Structure", y="Percentage (%)",
                     title="Large Intestine Anatomical Structures - Cell Types greater than 8% share, by Donor Sex and Analysis Tool") +
                scale_fill_manual(name="Cell Type",values=color_pal_kid2) +
                scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
                theme_light() +
                theme(legend.position = "bottom",
                      legend.title = element_text(face="bold", size=10),
                      legend.key.spacing.y = unit(.3,"pt"),
                      strip.background = element_rect(fill = "#cdcecf"),
                      strip.text = element_text(color="black", size=10, hjust=0)) +
                guides(fill=guide_legend(ncol=4))

# Export pdf
ggsave(file=paste0("../output/barplots/lrgIntestines_as_celltype_bargraph_0.2.pdf"),
       lrg_bg, width=14, height=12, units="in")

lrg_bg
```

#### Small Intestine - Anatomical Structure - Cell Types

```{r bar_smallIntestine_as_1, fig.height=12, fig.width=14, message=FALSE, warning=FALSE, paged.print=FALSE}

  
# Prepare for visualization
sml <- sml %>% 
  mutate(cell_lab = paste0(cell_id,": ",annotation_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) %>%
  ddply(.(sex,tool,as,cell_lab), summarise,
        obvs = length(cell_lab),
        cell_count = sum(cell_count),
        mn_pct = mean(percentage)) %>%
  arrange(sex, tool, as, desc(mn_pct)) %>%
  group_by(sex, tool, as) %>%
  mutate(rank = row_number()) %>%
  filter(mn_pct >= .08)

# Factor Organ AS
sml$as <- factor(sml$as, 
                 levels=c("Superior part of Duodenum", 
                          "Descending part of Duodenum",
                          "Horizontal part of Duodenum",
                          "Ascending part of Duodenum",
                          "Jejunum",
                          "Ileum",
                          "Distal part of Ileum"),
                 labels=c("Superior\nDuodenum", 
                          "Descending\nDuodenum",
                          "Horizontal\nDuodenum",
                          "Ascending\nDuodenum",
                          "Jejunum",
                          "Ileum",
                          "Distal\nIleum"))

# Factor Organ Cell Types
sml$cell_lab <- factor(sml$cell_lab,
                       levels=c("ASCTB-TEMP:b: B","ASCTB-TEMP:cd8-t: CD8+ T",
                                "ASCTB-TEMP:cycling-plasma-cell: Cycling plasma cell",
                                "ASCTB-TEMP:cycling-ta: Cycling TA",
                                "ASCTB-TEMP:endothelial: Endothelial",
                                "CL:1000334: Enterocyte Of Epithelium Of Small Intestine",
                                "ASCTB-TEMP:enterocyte: Enterocyte",
                                "CL:1000495: Small Intestine Goblet Cell",
                                "ASCTB-TEMP:goblet-cell: Goblet cell",
                                "ASCTB-TEMP:mesoderm-1-hand1-: Mesoderm 1 (HAND1+)",
                                "ASCTB-TEMP:microfold-cell: Microfold cell",
                                "ASCTB-TEMP:paneth: Paneth",
                                "ASCTB-TEMP:smooth-muscle: Smooth muscle",
                                "ASCTB-TEMP:stroma: Stroma"),
                       labels=c("ASCTB-TEMP: B",
                                "ASCTB-TEMP: CD8+ T",
                                "ASCTB-TEMP: Cycling Plasma Cell",
                                "ASCTB-TEMP: Cycling TA",
                                "ASCTB-TEMP: Endothelial",
                                "ASCTB-TEMP: Enterocyte",
                                "CL:1000334: Enterocyte of Epithelium",
                                "CL:1000495: Goblet Cell",
                                "ASCTB-TEMP: Goblet cell",
                                "ASCTB-TEMP: Mesoderm 1 (HAND1+)",
                                "ASCTB-TEMP: Microfold cell",
                                "ASCTB-TEMP: Paneth",
                                "ASCTB-TEMP: Smooth muscle",
                                "ASCTB-TEMP: Stroma"))

color_pal_kid <- c("#CD8490","#AADCDF","#88B1B3",
                   "#BAD2E7","#9AADBE","#7E8E9C","#626F7A",
                   "#D6B6D7","#AB92AC",
                   "#9C534E","#582FC5","#A1ACD2")

color_pal_kid2 <- c("#F9CE8D","#E97B74","#556E6F","#70A5A8","#A295C9","#EDB8AC","#CD8490",
                    "#E9E4F3","#D6B6D7","#AADCDF","#BAD2E7","#7495AF","#8DC599","#6F9A78",
                    "#9C534E","#C39A5D","#A17F4D","#675868")

sml_bg <- sml %>% arrange(sex,tool,as,mn_pct) %>%
              ggplot(aes(x=as,y=mn_pct)) +
                geom_bar(aes(fill=cell_lab), stat="identity") +
                facet_wrap(facets=vars(tool, sex), ncol=2) +
                labs(x="Anatomical Structure", y="Percentage (%)",
                     title="Small Intestine Anatomical Structures - Cell Types greater than 8% share, by Donor Sex and Analysis Tool") +
                scale_fill_manual(name="Cell Type",values=color_pal_kid2) +
                scale_y_continuous(expand = c(0,0),limits=c(0,1.05)) +
                theme_light() +
                theme(legend.position = "bottom",
                      legend.title = element_text(face="bold", size=10),
                      legend.key.spacing.y = unit(.3,"pt"),
                      strip.background = element_rect(fill = "#cdcecf"),
                      strip.text = element_text(color="black", size=10, hjust=0)) +
                guides(fill=guide_legend(ncol=4))

# Export pdf
ggsave(file=paste0("../output/barplots/smlIntestines_as_celltype_bargraph_0.2.pdf"),
       sml_bg, width=14, height=12, units="in")

sml_bg
```



```{r not visualized}
# Liver - 81 records
lvr <- df[df$organ==levels(df$organ)[5],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) 
lvr$as <- factor(lvr$as)
lvr$cell_lab <- factor(lvr$cell_lab)

# Breast - 14 records
brt <- df[df$organ==levels(df$organ)[3],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
brt$as <- factor(brt$as)
brt$cell_lab <- factor(brt$cell_lab)

# Prostate - 24 records
pro <- df[df$organ==levels(df$organ)[7],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
pro$as <- factor(pro$as)
pro$cell_lab <- factor(pro$cell_lab)

# Skin - 50 records
skn <- df[df$organ==levels(df$organ)[8],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
skn$as <- factor(skn$as)
skn$cell_lab <- factor(skn$cell_lab)

# Spleen - 75 records
spl <- df[df$organ==levels(df$organ)[10],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) 
spl$as <- factor(spl$as)
spl$cell_lab <- factor(spl$cell_lab)

# Ureter - 71 records
utr <- df[df$organ==levels(df$organ)[11],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex)
utr$as <- factor(utr$as)
utr$cell_lab <- factor(utr$cell_lab)

# Bladder - 26 records
bld <- df[df$organ==levels(df$organ)[12],] %>%
  mutate(cell_lab = paste0(cell_id,": ",cell_label)) %>%
  select(organ,as,cell_lab,cell_count,percentage,tool,sex) 
bld$as <- factor(bld$as)
bld$cell_lab <- factor(bld$cell_lab)

```





