---
title: "CT-Pop Biomarker Density Graphs Notebook"
output: html_document
---

# Environment 

```{r environment, message=FALSE, warning=FALSE}
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(magrittr)
library(ggplot2)
library(ggridges)

options(scipen = 999)
```

# Load Data

```{r load}
df <- read.csv("https://raw.githubusercontent.com/x-atlas-consortia/hra-pop/main/output-data/v0.10.2/reports/atlas/validation-b-mean-expression.csv",
               header=T, fileEncoding = "UTF8")
```

# Prepare data

```{r prep}
# Removes proteomics tools
df <- df[df$tool != "sc_proteomics",]

# Names alignment
names(df)[4] <- "organ"
names(df)[7] <- "cell"
names(df)[8] <- "gene"

# Prepare organ labels
df$organ <- df$organ %>% 
  str_replace(" of body", "") %>% 
  str_to_title()

# Factor organs
df$organ <- factor(df$organ)

# Prepare Cell Labels
df$cell <- 
  df$cell %>%
  str_to_upper()

# Prepare tool
df$tool <- df$tool %>%
  str_to_title() 

# Factor tools
df$tool <- factor(df$tool)
```

## Rank biomarkers and subsetting data 

```{r ranking}
# Count the number cells with a biomarker measurements for each organ
df <- df %>% 
  arrange(organ, gene) %>%
  select(tool, organ, gene, cell, mean_gene_expression_value) %>%
  group_by(organ, gene) %>%
  mutate(ct=row_number())

# Ranks top 20 biomarkers based on the number measurements for each organ 
# - by tool
gene_ranks_o_t <- df %>%
  ddply(.(tool, organ, gene), summarise,
        max_cells = max(ct)) %>%
  arrange(tool, organ, desc(max_cells)) %>%
  group_by(tool, organ) %>%
  mutate(rank=row_number()) %>%
  filter(rank < 21)
 
# Overall
gene_ranks_o <- df %>%
  ddply(.(organ, gene), summarise,
        max_cells = max(ct)) %>%
  arrange(organ, desc(max_cells)) %>%
  group_by(organ) %>%
  mutate(rank=row_number()) %>%
  filter(rank < 21)

# Selects all relevant measurements for top 20 biomarker measurements by organ
# Overall
gene_sets <- right_join(df[,2:6], gene_ranks_o,by=c("organ","gene")) %>%
  arrange(organ, rank)

# Tools
gene_sets_t <- right_join(df, gene_ranks_o_t,by=c("tool","organ","gene")) %>%
  arrange(tool, organ, rank)
```

## Ranking - Overall
```{r overall-ranks}
# Identify top 40 biomarkers across all data 
gene_ranks_f <- 
  unique(gene_sets[,c(1,2,6)]) %>%
  group_by(organ) %>%
  mutate(rank_organ = row_number()) %>%
  ungroup() %>%
  arrange(desc(max_cells)) %>%
  mutate(rank_overall = row_number())

# Selects all relevant measurements for top 40 biomarkers across all data.
gene_ranks_f1 <- gene_ranks_f[gene_ranks_f$rank_overall < 41,]

## Joins gene sets data with overall ranks 
gene_sets_o1 <- 
  right_join(gene_sets, 
             gene_ranks_f1[,c(1,2,4,5)],
             by=c("organ","gene")) %>%
  arrange(rank_overall)

## Factoring variables
gene_sets_o1$organ <- 
  factor(as.character(gene_sets_o1$organ),
         levels=c("Heart","Lung","Kidney","Small Intestine","Large Intestine"))
gene_sets_o1$gene <- 
  factor(gene_sets_o1$gene)

## Selects all data of top 20 biomarkers for each organ
## Joins gene sets data with organ gene ranks 
gene_sets_o2 <- 
  gene_sets %>%
  select(organ, gene, max_cells) %>%
  distinct() %>%
  arrange(organ, desc(max_cells)) %>%
  group_by(organ) %>%
  mutate(rank_overall = row_number()) %>% 
  filter(rank_overall <= 20)

gene_sets_o2 <- right_join(gene_sets, 
                           gene_sets_o2[,c(1,2,4)], 
                           by=c("organ","gene")) %>% 
                  distinct() %>%
                  arrange(organ,gene,rank_overall)

## Factoring variables
gene_sets_o2$organ <- 
  factor(as.character(gene_sets_o2$organ),
         levels=c("Heart","Lung","Kidney","Small Intestine","Large Intestine"))
```

# Visualization

## Overall - Top 20 Biomarkers (cell counts) - Gene Expression Density Ridge Plots

```{r densityplot_biomarkers_organ_overall, fig.height=10, fig.width=6, message=FALSE, warning=FALSE}
theme_set(theme_minimal())

# Factor gene + cell observation label
labs <- unique(gene_sets_o1[order(gene_sets_o1$max_cells),c(2,6)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") 
gene_sets_o1 <- left_join(gene_sets_o1,labs,by=c("gene","max_cells")) 
gene_sets_o1$facs <- factor(gene_sets_o1$facs, 
                            levels=labs$facs)


# Visualization
color_pal <- c("#EF5350","#FFD54F","#9575CD","#0288D1","#4CAF50")
                  
overall_dr <- gene_sets_o1 %>%
              ggplot(aes(x=log(mean_gene_expression_value), y=facs)) +
                  geom_density_ridges(aes(fill=organ),
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.55) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)") +
                  scale_fill_manual(name="Organ", values=color_pal) +
                  theme(legend.justification = "top")

# Export visualization
ggsave(file=paste0("../output/densityplots/overall_top20_gene_expression_v0.1.pdf"),
       overall_dr, width=5, height=6.5, units="in")
ggsave(file=paste0("../output/densityplots/overall_top20_gene_expression_v0.1.png"),
       overall_dr, width=5, height=6.5, units="in")

overall_dr
```


## Heart 

### Overall - Top 20 Biomarkers (cell counts) - Gene Expression Density Ridge Plots

```{r density_heart, fig.height=6, fig.width=4.75, message=FALSE, warning=FALSE}
theme_set(theme_minimal())

# Subset organ
heart <- gene_sets_o2[gene_sets_o2$organ=="Heart",] 

# Unique factor with cell count
labs <- unique(heart[order(heart$max_cells),c(2,6)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") 
heart <- left_join(heart,labs[,c(1,3)],by="gene")
heart$facs <- factor(heart$facs, 
                     levels=labs$facs)

# Color palette 
color_org <- c("#EF5350")

# Visualization
hrt_dr <- heart %>% 
            filter(facs != "NA") %>%
            ggplot(aes(x=log(mean_gene_expression_value), 
                       y=facs)) +
              geom_density_ridges(aes(fill=organ), 
                                  na.rm = TRUE,,
                                  bandwidth = 0.15,
                                  rel_min_height = 0.01, alpha=.85) +
              labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)") +
              scale_fill_manual(name="Organ", values=color_org) +
              theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_v0.1.pdf"),
       hrt_dr, width=4.75, height=6, units="in")
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_v0.1.png"),
       hrt_dr, width=4.75, height=6, units="in")
hrt_dr
```

### By Tools

```{r ds_hrt_tools, fig.height=5, fig.width=5}
# Select Data
heart_tools <- gene_sets_t %>% 
  filter(organ == "Heart") %>%
  mutate(tool = factor(tool, 
                       levels=c("Azimuth","Celltypist","Popv"),
                       labels=c("Azimuth","CellTypist","PopV")))

# Unique factor with cell count
labs <- unique(heart_tools[order(heart_tools$max_cells),c(1,3,7)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") %>%
  arrange(tool, desc(max_cells))
heart_tools <- left_join(heart_tools,labs[,c(1,2,4)],by=c("tool","gene"))
heart_tools$facs <- factor(heart_tools$facs, 
                           levels=labs$facs)

# Color palette 
color_org <- c("#EF5350","#E57373","#EF9A9A")

# Visualization
# Azimuth
hrt_dr_az <- heart_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "Azimuth") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="Azimuth") +
                  scale_fill_manual(name="Organ", values=color_org[1]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_AZ_v0.1.pdf"),
       hrt_dr_az, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_AZ_v0.1.png"),
       hrt_dr_az, width=5, height=5, units="in")

# CellTypist 
hrt_dr_ct <- heart_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "CellTypist") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="CellTypist") +
                  scale_fill_manual(name="Organ", values=color_org[2]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_CT_v0.1.pdf"),
       hrt_dr_ct, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_CT_v0.1.png"),
       hrt_dr_ct, width=5, height=5, units="in")

# PopV
hrt_dr_pv <- heart_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "PopV") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="PopV") +
                  scale_fill_manual(name="Organ", values=color_org[3]) +
                  theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_PV_v0.1.pdf"),
       hrt_dr_pv, width=4, height=5, units="in")
ggsave(file=paste0("../output/densityplots/heart_top20_gene_expression_PV_v0.1.png"),
       hrt_dr_pv, width=4, height=5, units="in")

hrt_dr_az
hrt_dr_ct
hrt_dr_pv
```


## Lung 

### Overall - Top 20 Biomarkers (cell counts) - Gene Expression Density Ridge Plots
```{r density_lung, fig.height=6, fig.width=4.75, message=FALSE, warning=FALSE}
theme_set(theme_minimal())

# Subset organ
lung <- gene_sets_o2[gene_sets_o2$organ=="Lung",] 

# Unique factor with cell count
labs <- unique(lung[order(lung$max_cells),c(2,6)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") 
lung <- left_join(lung,labs[,c(1,3)],by="gene")
lung$facs <- factor(lung$facs, 
                    levels=labs$facs)

# Color palette 
color_org <- c("#FFD54F")

# Visualization
lung_dr <- lung %>% 
            filter(facs != "NA") %>%
            ggplot(aes(x=log(mean_gene_expression_value), 
                       y=facs)) +
              geom_density_ridges(aes(fill=organ), 
                                  na.rm = TRUE,,
                                  bandwidth = 0.15,
                                  rel_min_height = 0.01, alpha=.85) +
              labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)") +
              scale_fill_manual(name="Organ", values=color_org) +
              theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_v0.1.pdf"),
       lung_dr, width=4.75, height=6, units="in")
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_v0.1.png"),
       lung_dr, width=4.75, height=6, units="in")
lung_dr
```

### By Tools

```{r ds_lng_tools, fig.height=5, fig.width=5}
# Select Data
lung_tools <- gene_sets_t %>% 
  filter(organ == "Lung") %>%
  mutate(tool = factor(tool, 
                       levels=c("Azimuth","Celltypist","Popv"),
                       labels=c("Azimuth","CellTypist","PopV")))

# Unique factor with cell count
labs <- unique(lung_tools[order(lung_tools$max_cells),c(1,3,7)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") %>%
  arrange(tool, desc(max_cells))
lung_tools <- left_join(lung_tools,labs[,c(1,2,4)],by=c("tool","gene"))
lung_tools$facs <- factor(lung_tools$facs, 
                           levels=labs$facs)

# Color palette 
color_org <- c("#FFB300","#FFCA28","#FFE082")

# Visualization
# Azimuth
lng_dr_az <- lung_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "Azimuth") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="Azimuth") +
                  scale_fill_manual(name="Organ", values=color_org[1]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_AZ_v0.1.pdf"),
       lng_dr_az, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_AZ_v0.1.png"),
       lng_dr_az, width=5, height=5, units="in")

# CellTypist 
lng_dr_ct <- lung_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "CellTypist") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="CellTypist") +
                  scale_fill_manual(name="Organ", values=color_org[2]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_CT_v0.1.pdf"),
       lng_dr_ct, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_CT_v0.1.png"),
       lng_dr_ct, width=5, height=5, units="in")

# PopV
lng_dr_pv <- lung_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "PopV") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="PopV") +
                  scale_fill_manual(name="Organ", values=color_org[3]) +
                  theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_PV_v0.1.pdf"),
       lng_dr_pv, width=4, height=5, units="in")
ggsave(file=paste0("../output/densityplots/lung_top20_gene_expression_PV_v0.1.png"),
       lng_dr_pv, width=4, height=5, units="in")

lng_dr_az
lng_dr_ct
lng_dr_pv
```




## Kidney 

### Overall - Top 20 Biomarkers (cell counts) - Gene Expression Density Ridge Plots
```{r density_kidney, fig.height=6, fig.width=4.75, message=FALSE, warning=FALSE}
# Subset organ
kid <- gene_sets_o2[gene_sets_o2$organ=="Kidney",] 

# Unique factor with cell count
labs <- unique(kid[order(kid$max_cells),c(2,6)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") 
kid <- left_join(kid,labs[,c(1,3)],by="gene")
kid$facs <- factor(kid$facs, 
                   levels=labs$facs)

# Color palette 
color_org <- c("#9575CD")
#"#7495AF","#8DC599"

# Visualization
kid_dr <- kid %>% 
            filter(facs != "NA") %>%
            ggplot(aes(x=log(mean_gene_expression_value), 
                       y=facs)) +
              geom_density_ridges(aes(fill=organ), 
                                  na.rm = TRUE,,
                                  bandwidth = 0.15,
                                  rel_min_height = 0.01, alpha=.85) +
              labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)") +
              scale_fill_manual(name="Organ", values=color_org) +
              theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/kidney_top20_gene_expression_v0.1.pdf"),
       kid_dr, width=4.75, height=6, units="in")
ggsave(file=paste0("../output/densityplots/kidney_top20_gene_expression_v0.1.png"),
       kid_dr, width=4.75, height=6, units="in")
kid_dr
```

### By Tools

```{r ds_kid_tools, fig.height=5, fig.width=5}
# Select Data
kidney_tools <- gene_sets_t %>% 
  filter(organ == "Kidney") %>%
  mutate(tool = factor(tool, 
                       levels=c("Azimuth"),
                       labels=c("Azimuth")))

# Unique factor with cell count
labs <- unique(kidney_tools[order(kidney_tools$max_cells),c(1,3,7)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") %>%
  arrange(tool, desc(max_cells))
kidney_tools <- left_join(kidney_tools,labs[,c(1,2,4)],by=c("tool","gene"))
kidney_tools$facs <- factor(kidney_tools$facs, 
                           levels=labs$facs)

# Color palette 
color_org <- c("#B39DDB")

# Visualization
# Azimuth
kid_dr_az <- kidney_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "Azimuth") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="Azimuth") +
                  scale_fill_manual(name="Organ", values=color_org[1]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/kidney_top20_gene_expression_AZ_v0.1.pdf"),
       kid_dr_az, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/kidney_top20_gene_expression_AZ_v0.1.png"),
       kid_dr_az, width=5, height=5, units="in")

kid_dr_az
```

## Small Intestine 

### Overall - Top 20 Biomarkers (cell counts) - Gene Expression Density Ridge Plots
```{r density_sml, fig.height=6, fig.width=4.75, message=FALSE, warning=FALSE}
# Subset organ
sml <- gene_sets_o2[gene_sets_o2$organ=="Small Intestine",] 

# Unique factor with cell count
labs <- unique(sml[order(sml$max_cells),c(2,6)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") 
sml <- left_join(sml,labs[,c(1,3)],by="gene")
sml$facs <- factor(sml$facs, 
                   levels=labs$facs)

# Color palette 
color_org <- c("#0288D1")

# Visualization
sml_dr <- sml %>% 
            filter(facs != "NA") %>%
            ggplot(aes(x=log(mean_gene_expression_value), 
                       y=facs)) +
              geom_density_ridges(aes(fill=organ), 
                                  na.rm = TRUE,,
                                  bandwidth = 0.15,
                                  rel_min_height = 0.01, alpha=.85) +
              labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)") +
              scale_fill_manual(name="Organ", values=color_org) +
              theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/smallIntestine_top20_gene_expression_v0.1.pdf"),
       sml_dr, width=4.75, height=6, units="in")
ggsave(file=paste0("../output/densityplots/smallIntestine_top20_gene_expression_v0.1.png"),
       sml_dr, width=4.75, height=6, units="in")
sml_dr
```
### By Tools

```{r ds_sml_tools, fig.height=5, fig.width=5}
# Select Data
sml_tools <- gene_sets_t %>% 
  filter(organ == "Small Intestine") %>%
  mutate(tool = factor(tool, 
                       levels=c("Celltypist","Popv"),
                       labels=c("CellTypist","PopV")))

# Unique factor with cell count
labs <- unique(sml_tools[order(sml_tools$max_cells),c(1,3,7)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") %>%
  arrange(tool, desc(max_cells))
sml_tools <- left_join(sml_tools,labs[,c(1,2,4)],by=c("tool","gene"))
sml_tools$facs <- factor(sml_tools$facs, 
                         levels=labs$facs)

# Color palette 
color_org <- c("#4FC3F7","#B3E5FC")

# Visualization
# CellTypist 
sml_dr_ct <- sml_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "CellTypist") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="CellTypist") +
                  scale_fill_manual(name="Organ", values=color_org[1]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/smallIntestine_top20_gene_expression_CT_v0.1.pdf"),
       sml_dr_ct, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/smallIntestine_top20_gene_expression_CT_v0.1.png"),
       sml_dr_ct, width=5, height=5, units="in")

# PopV
sml_dr_pv <- sml_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "PopV") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="PopV") +
                  scale_fill_manual(name="Organ", values=color_org[2]) +
                  theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/smallIntestine_top20_gene_expression_PV_v0.1.pdf"),
       sml_dr_pv, width=4, height=5, units="in")
ggsave(file=paste0("../output/densityplots/smallIntestine_top20_gene_expression_PV_v0.1.png"),
       sml_dr_pv, width=4, height=5, units="in")

sml_dr_ct
sml_dr_pv
```


## Large Intestine

### Overall - Top 20 Biomarkers (cell counts) - Gene Expression Density Ridge Plots
```{r density_lrg, fig.height=6, fig.width=4.75, message=FALSE, warning=FALSE}
# Subset organ
lrg <- gene_sets_o2[gene_sets_o2$organ=="Large Intestine",] 

# Unique factor with cell count
labs <- unique(lrg[order(lrg$max_cells),c(2,6)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") 
lrg <- left_join(lrg,labs[,c(1,3)],by="gene")
lrg$facs <- factor(lrg$facs, 
                   levels=labs$facs)

# Color palette 
color_org <- c("#4CAF50")

# Visualization
lrg_dr  <- lrg %>% 
            filter(facs != "NA") %>%
            ggplot(aes(x=log(mean_gene_expression_value), 
                       y=facs)) +
              geom_density_ridges(aes(fill=organ), 
                                  na.rm = TRUE,,
                                  bandwidth = 0.15,
                                  rel_min_height = 0.01, alpha=.85) +
              labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)") +
              scale_fill_manual(name="Organ", values=color_org) +
              theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/largeIntestine_top20_gene_expression_v0.1.pdf"),
       lrg_dr, width=4.75, height=6, units="in")
ggsave(file=paste0("../output/densityplots/largeIntestine_top20_gene_expression_v0.1.png"),
       lrg_dr, width=4.75, height=6, units="in")
lrg_dr
```


```{r ds_lrg_tools, fig.height=5, fig.width=5}
# Select Data
lrgI_tools <- gene_sets_t %>% 
  filter(organ == "Large Intestine") %>%
  mutate(tool = factor(tool, 
                       levels=c("Celltypist","Popv"),
                       labels=c("CellTypist","PopV")))

# Unique factor with cell count
labs <- unique(lrgI_tools[order(lrgI_tools$max_cells),c(1,3,7)]) %>%
  mutate(facs = paste0(gene," (",max_cells,")")) %>%
  filter(gene != "NA") %>%
  arrange(tool, desc(max_cells))
lrgI_tools <- left_join(lrgI_tools,labs[,c(1,2,4)],by=c("tool","gene"))
lrgI_tools$facs <- factor(lrgI_tools$facs, 
                           levels=labs$facs)

# Color palette 
color_org <- c("#66BB6A","#A5D6A7")

# Visualization
# CellTypist 
lrg_dr_ct <- lrgI_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "CellTypist") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="CellTypist") +
                  scale_fill_manual(name="Organ", values=color_org[1]) +
                  theme(legend.position = "none")
# Export pdf
ggsave(file=paste0("../output/densityplots/largeIntestine_top20_gene_expression_CT_v0.1.pdf"),
       lrg_dr_ct, width=5, height=5, units="in")
ggsave(file=paste0("../output/densityplots/largeIntestine_top20_gene_expression_CT_v0.1.png"),
       lrg_dr_ct, width=5, height=5, units="in")

# PopV
lrg_dr_pv <- lrgI_tools %>% 
                filter(facs != "NA") %>%
                filter(tool == "PopV") %>%
                ggplot(aes(x=log(mean_gene_expression_value), 
                           y=facs)) +
                  geom_density_ridges(aes(fill=organ), 
                                      na.rm = TRUE,,
                                      bandwidth = 0.15,
                                      rel_min_height = 0.01, alpha=.85) +
                  labs(x="Mean Gene Expression Values (log2)", y="Biomarker (Cells)",
                       title="PopV") +
                  scale_fill_manual(name="Organ", values=color_org[2]) +
                  theme(legend.position = "none")

# Export pdf
ggsave(file=paste0("../output/densityplots/largeIntestine_top20_gene_expression_PV_v0.1.pdf"),
       lrg_dr_pv, width=4, height=5, units="in")
ggsave(file=paste0("../output/densityplots/largeIntestine_top20_gene_expression_PV_v0.1.png"),
       lrg_dr_pv, width=4, height=5, units="in")

lrg_dr_ct
lrg_dr_pv
```