---
title: "HRApop Sankey Universe"
output:
  html_document:
    df_print: paged
---

# Environment

```{r message=FALSE, warning=FALSE}

#libraries
library(tidyverse)
library(scales) #for scatter graph and bar graph
library(ggrepel) # to jitter labels
library(networkD3) #for Sankey
library(RColorBrewer) # for plots
library(ggcorrplot) # for tissue block similarity matrix
library(lsa) # for tissue block similarity matrix
library(rmarkdown) # for rendering this

#load themes
source("scripts/themes.R")

# global variables
hra_pop_version = "0.10.2"

# set up color palettes
# extend Brewer color palettes
nb.cols <- 16
cat_colors <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)
```

We will use `networkD3` to render the Sankey diagram. On the data side, we need two tibbles:\
 * NODES with NodeId\
 * LINKS with Source, Target, Value

# Supplemental Fig. Sankey (universe)

## Load Data

```{r message=FALSE, warning=FALSE}
# atlas data only
sankey_universe = read_csv(paste("../../hra-pop/output-data/v",hra_pop_version,"/reports/universe-ad-hoc/sankey.csv", sep = ""))
```

## Create bins for donor age and BMI

### Define function to create bins
```{r}
add_bin = function(original_column, bins, labels){
  # Create new column with bins
  result <- cut(original_column, breaks = bins, labels = labels, include.lowest = TRUE)
  
  return(as.character(result))
}
```

### Add bins
```{r message=FALSE}

# create new tibble
sankey_universe_with_bins = sankey_universe 

# Create bins and labels
bins_age = c(0, 18, 35, 50, 65, 85, 120)  # Define bin edges
labels_age = c("Underage", "Young adult", "Adult", "Middle", "Senior", "Very senior") 
bins_bmi = c(0,18.5, 24.9, 29.9, 30.0)
labels_bmi = c("underweight", "healthy", "overweight", "obese")

sankey_universe_with_bins$donor_age_binned = add_bin(sankey_universe_with_bins$donor_age, bins_age, labels_age)
sankey_universe_with_bins$donor_bmi_binned = add_bin(sankey_universe_with_bins$donor_bmi, bins_bmi, labels_bmi)


```

## Create subset and replace NAs, unify unknown values, adjust portal values
```{r}
subset_sankey = sankey_universe_with_bins %>% 
  select(portal, donor_sex, organ_name, cell_type_annotation_tool, donor_race, donor_bmi_binned, donor_age_binned, is_rui_registered, is_atlas_dataset) %>% 
  
  # replace NAs
  replace_na(list(donor_sex = "Unknown Sex")) %>% 
  replace_na(list(donor_race = "Unknown Race")) %>% 
  replace_na(list(organ_name = "Unknown Organ")) %>% 
  replace_na(list(cell_type_annotation_tool = "No Cell Summary")) %>%
  
  # replace NAs for binned variables
  replace_na(list(donor_bmi_binned = "Unknown BMI")) %>%
  replace_na(list(donor_age_binned = "Unknown Age")) %>%
  # unify left and right kidney
  mutate(organ_name = ifelse(organ_name == "left kidney" | organ_name == "right kidney", "kidney", organ_name))
  
  # unify unknown values in race and sex
  subset_sankey$donor_race[subset_sankey$donor_race == "Unknown"] <- "Unknown Race"
  subset_sankey$donor_sex[subset_sankey$donor_sex == "Unknown"] <- "Unknown Sex"
  
  # replace HCA and LungMAP with CxG
  subset_sankey$portal[subset_sankey$portal == "HCA"] <- "CxG"
  subset_sankey$portal[subset_sankey$portal == "NHLBI/LungMap"] <- "CxG"
  
  # turn LGL into meaningful CHAR for rui and atlas
  subset_sankey$is_rui_registered[subset_sankey$is_rui_registered == TRUE] <- "Is RUI-registered"
  subset_sankey$is_rui_registered[subset_sankey$is_rui_registered == FALSE] <- "Is not RUI-registered"
  subset_sankey$is_atlas_dataset[subset_sankey$is_atlas_dataset == TRUE] <- "Is Atlas Dataset"
  subset_sankey$is_atlas_dataset[subset_sankey$is_atlas_dataset == FALSE] <- "Is not Atlas Dataset"


```

## Create nodes
```{r}
p = subset_sankey %>% 
  group_by(portal) %>% summarize()

d = subset_sankey %>% 
  group_by(donor_sex) %>% summarize()

a = subset_sankey %>% 
  group_by(donor_age_binned) %>% summarize()

b = subset_sankey %>%
  group_by(donor_bmi_binned) %>% summarize()

r = subset_sankey %>% 
  group_by(donor_race) %>% summarize()

o = subset_sankey %>% 
  group_by(organ_name) %>% summarize()

c = subset_sankey %>% 
  group_by(cell_type_annotation_tool) %>% summarize()

rui = subset_sankey %>% 
  group_by(is_rui_registered) %>% summarize()

atlas = subset_sankey %>% 
  group_by(is_atlas_dataset) %>% summarize()

unique_name=list()
unique_name = unlist(append(unique_name, c(p, d, a, b, r, o, c, rui, atlas)))
unique_name = list(unique_name)

nodes = as.data.frame(tibble(name = character()))
```

## Create links
 Use this order: 
  * sources
  * donor sex
  * donor age
  * donor BMI
  * donor ethnicity
  * organ
  * cell type annotation tool used
  * RUI registered or not
  * ‘atlas’ dataset or not 
```{r}

for(u in unique_name){
  nodes = nodes %>% 
    add_row(name=u)
}

nodes$index <- 1:nrow (nodes) 
nodes

nodes$index = nodes$index-1
nodes

portal_sex = subset_sankey %>% 
  group_by(portal, donor_sex) %>% 
  summarize(count=n()) %>% 
  rename(
    source = portal,
    target = donor_sex,
    value=count
  )

sex_age = subset_sankey %>% 
  group_by(donor_sex, donor_age_binned) %>% 
  summarize(count=n()) %>% 
  rename(
    source = donor_sex,
    target = donor_age_binned,
    value=count
  )

age_bmi = subset_sankey %>% 
  group_by(donor_age_binned, donor_bmi_binned) %>% 
  summarize(count=n()) %>% 
  rename(
    source = donor_age_binned,
    target = donor_bmi_binned,
    value=count
  )

bmi_race = subset_sankey %>% 
  group_by(donor_bmi_binned, donor_race) %>% 
  summarize(count=n()) %>% 
  rename(
    source = donor_bmi_binned,
    target = donor_race,
    value=count
  )

race_organ = subset_sankey %>%
  group_by(donor_race, organ_name) %>%
  summarize(count=n()) %>%
  rename(
    source = donor_race,
    target = organ_name,
    value=count
  )

organ_ctann = subset_sankey %>% 
  group_by(organ_name, cell_type_annotation_tool) %>% 
  summarize(count=n()) %>% 
  rename(
    source = organ_name,
    target = cell_type_annotation_tool,
    value=count
  )

ctann_rui = subset_sankey %>% 
  group_by(cell_type_annotation_tool, is_rui_registered) %>% 
  summarize(count=n()) %>% 
  rename(
    source = cell_type_annotation_tool,
    target = is_rui_registered,
    value=count
  )

rui_atlas = subset_sankey %>% 
  group_by(is_rui_registered, is_atlas_dataset) %>% 
  summarize(count=n()) %>% 
  rename(
    source = is_rui_registered,
    target = is_atlas_dataset,
    value=count
  )


prep_links = as.data.frame(bind_rows(portal_sex, sex_age, age_bmi, bmi_race, race_organ, organ_ctann, ctann_rui, rui_atlas))
prep_links 

links = prep_links 
```


## Rename node and link tables
```{r}
names(nodes)[1] = "source"
prep_links = left_join(prep_links, nodes,by="source")

names(nodes)[1] = "target"
prep_links = left_join(prep_links, nodes,by="target")
prep_links

prep_links = prep_links[,c(4,5,3)]
names(prep_links)[1:2] = c("source", "target")
names(nodes)[1] = "name"
```

## Draw the Sankey diagram

```{r}
p <- sankeyNetwork(Links = prep_links, Nodes = nodes, Source = "source",
                   Target = "target", Value = "value", NodeID = "name", 
                   units = "occurrences", fontSize = 15, nodeWidth = 30, width=1200, height = 650)

p
```

## Export the Sankey diagram as HTML

```{r}
saveNetwork(p, "../docs/sankey_universe.html")
```
